Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2022-11-21 latex/comp_or.tex
Paper update

@@

In the scenario of aerial pickup and delivery of goods in a distribution network, transport aviation faces risks of cargo unbalancing due to the urgency required for loading for rapid take-off and mission accomplishment, especially in times of crisis, public calamity, client contract short lead times, or any external pressure for immediate take-off. Also, there is no technological assistance to help the load and trip planners with the huge number of demands for transport in each hub.
This enables other risks such as improper delivery, excessive fuel burn, and longer than necessary turn-around time. We contributed by modelling and solving a new problem of planning the loading and routing of an aircraft according to a utility score, weight and balance principles, and fuel consumption in a tour of simultaneous pickup and delivery at intermediate hubs.
This new hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem (ACLP+RPDP)}, is mathematically modelled using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.
We also contributed by carrying out multiple experiments with the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history.
These challenging benchmark instances are made publicly available.
We also created a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as { \color{red} it was the unique method that managed to solve all testing scenarios.}

% Reescrever

@@

Air cargo transport involves several sub-problems that are difficult to solve. Recently, \cite[p. 401]{BrandtStefan2019} defined the {\it Air Cargo Load Planning Problem} (ACLPP) as four sub-problems: {\it Aircraft Configuration Problem} (ACP), {\it Build-up Scheduling Problem} (BSP), {\it Air Cargo Palletization Problem} (APP) and {\it Weight and Balance Problem} (WBP). Several aspects were considered in this modelling: characteristics of the items to be transported (dimensions, scores, dangerousness, etc.); types and quantities of {\it unit load devices} (ULDs), commonly called pallets; when these pallets are assembled; how items are allocated to pallets; in which positions these pallets are to be placed; how the total cargo weight is balanced; etc. They also presented a comprehensive bibliographic survey of solving methods that had been developed in different situations.

However, there are still other important challenges in air cargo transport that go beyond the definition of the ACLPP, especially with regard to the flight itinerary and the loading and unloading at each destination (or node) of this travel plan. In this context, at least two more important sub-problems can be considered: pickup and delivery operations at each node, called {\it Simultaneous Pickup and Delivery Problem} (SPDP), and the search for the most profitable route, which is the well-known {\it Traveling Salesman Problem} (TSP), similar to the proposed by \cite{kaspi2019}, where the profit per unit time is defined as the profit gained after completing the tour divided by the total time required to complete the tour.

\cite{MesquitaSanches2023} explored these problems by solving them in an integrated form.

@@

		\cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$&  $.$ \\
		\cite{PaquaySchynsLimbourg2016,PaquayLimbourgSchynsOliveira2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$&  $.$ \\
		\cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$&  $.$ \\
		\cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$&  $.$  \\
		\cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$&  $.$  \\
		\cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$&  $.$ \\
		\cite{MesquitaSanches2023}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$&  $.$  \\
		{\bf This work}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$&  $\bigstar$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

This work proposes a method that attends these objectives: a pallet building and arrangement plan with a routing option that maximizes the benefit-cost ratio for the smooth execution of pickup and delivery transport missions. We develop a heuristic that can be executed on a simple handheld computer (like a laptop or a tablet) and that provides a solution quickly enough to keep this cargo handling time under an hour. This heuristic will reduce the stress that transport planners are subjected to, because they have to deal with a lot of information in planning the aircraft route, assembling the pallets, and picking up and delivering at each node. To the best of our knowledge, this is the first time that an air cargo transport problem that simultaneously involves APP, WBP, PDP and TSP has been addressed. This new problem is named {\it Air Cargo Load Planning with Routing, Pickup and Delivery Problem} (ACLP+RPDP).

@@

\cite{MongeauBes2003} presented a method based on linear integer programming to solve the problem of choosing and positioning containers on the {\it Airbus 340-300}. Safety and stability restrictions were considered, with the objective of minimizing fuel consumption.

@@

As already mentioned, \cite{LurkinSchyns2015} was the first work that simultaneously modelled WBP and SPDP in air cargo transport. The authors demonstrated that this problem is NP-hard and performed some experiments with real data, noting that their model offers better results than those obtained manually.

@@

\cite{PaquaySchynsLimbourg2016} presented a mathematical modelling to optimize the loading of heterogeneous 3D boxes on pallets with a truncated parallelepipeds format. Its objective is to maximise the volume used in containers, considering load balancing restrictions, the presence of fragile items and the possibility of rotating these boxes. \cite{PaquayLimbourgSchynsOliveira2018} developed some heuristics to solve this problem.

@@

\cite{wong2020}A developed a mathematical model and a tool based on mixed integer programming for optimizing cargo in aircraft with different pallet configurations. Balance restrictions and the presence of dangerous items were considered. \cite{eugene2021} integrated this tool to a digital simulation model, with a visualization and validation system, based on sensors that alert about load deviations.

@@

\cite{MesquitaSanches2023} contributed with a complex model and elaborate heuristics for the ACLP-RPDP that simultaneously solve 4 intractable sub-problems: APP, WBP, SPDP and TSP. They also compare the performances of four well known heuristics: Ant Colony Optimization, the Noising Method Optimization, the Greedy Randomized Adaptive Search Procedure, and Tabu Search.

As can be seen, except for \cite{MesquitaSanches2023}, the remaining works do not address air cargo palletization and load balancing with route optimization in a multi-leg flight plan, and none of them employ parallelization into their algorithms. This is the objective of our work: to reshape \cite{MesquitaSanches2023} solution process and algorithms to accommodate parallel features which can improve solution quality and performance.

@@

@@

Throughout this text, we call a {\it consolidated item} a set of items of the same destination stacked on a pallet and covered with a restraining net. It is considered unique, having the same attributes of its components, whose values are the sum of individual scores, weights and volumes. See Figure \ref{fig:larger2}. To ensure accuracy in pickup and delivery operations, consolidated items must remain on board until their destination.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Images/large.png}
	\caption{Consolidated items on pallets inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: https://www.sacprogram.org/en/Pages/Boeing-C-17-Globemaster-III.aspx}
	\label{fig:larger2}
\end{figure}

@@

We consider real scenarios with a smaller or a larger aircraft with payloads of $26,000 kg$ or $75,000 kg$ respectively. Both layouts are represented in Figures \ref{fig:smaller} and \ref{fig:larger}, where the pallets are identified by $p_i$.

@@

In both cases, the torque applied to the aircraft must keep its CG in the operational range, which corresponds to a percentage of the {\it Mean Aerodynamic Chord} \footnote{Chord is the distance between the leading and trailing edges of the wing, measured parallel to the normal airflow over the wing. The average length of the chord is known as the {\it Mean Aerodynamic Chord} (MAC).}: $0.556m$ in the smaller aircraft and $1.17m$ in the larger one. See Figure \ref{fig:lateral}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.22]{Images/lateral.png}
	\caption{Aircraft longitudinal cut showing red lines as pallets}
	%\small\textsuperscript{Source: the author}
	\label{fig:lateral}
\end{figure}

@@

We also make the following assumptions:
\begin{itemize}
	\item on each pallet, the items are distributed in such a way that their CG coincides with the centroid of the pallet;
	\item the CG of the payload must be at a maximum longitudinal distance of $limit^{CG}_{long}$ from the CG of the aircraft;
	\item in the larger aircraft, the CG of the payload must be at a maximum lateral distance of $limit^{CG}_{lat}$ from the CG of the aircraft;
	\item in the larger aircraft, pallets are distributed in two identical rows (with odd and even indices, respectively), and their centroids are at a distance $d^{CG}_{pallet}$ from the center-line of the aircraft.
\end{itemize}

@@

\bgroup
\def\arraystretch{1.2}
\begin{table}[H]
	\centering
	\small
	\begin{tabular}{MT}
		&Informally, ACLP+RPDP can be summarized as follows:\\
		\midrule
		max &  (items score sum) / (tour cost) of picked up and delivered items on a tour  \\
		\midrule
s.t.    & Along a tour, the set of unvisited nodes is updated. \\
		& In each node, an item may be included in at most one pallet.\\		
		& In each node, consolidated items are composed of items with the same destination. \\
		& Weight, volume and score of a consolidated item are the corresponding sum of their components.\\
		& Consolidated items remain on board until their destinations.\\	
		& Consolidated items can only be included in the same pallet if their destinations are the same.\\
		& Only items destined for the remaining nodes can be loaded.  \\
		& The lateral and longitudinal torques must be within the operational range of the aircraft.\\
		& Weight and volume limitations of pallets must be respected.\\
		& The total weight must be less than the aircraft payload or the total pallet capacity, whichever is the lowest.\\	
		\midrule
	\end{tabular}
	\normalsize
\end{table}
\egroup

@@

Our strategy is based on the fact that, in real cases, $K$\/ is usually small. Specifically, we will consider $K \leq 6$\/ throughout this work, which is a higher value than usual in {\it Brazilian Air Force} missions. As a result, if we have fast node-by-node solutions that allow us to construct a complete tour, we will be able to test all possible $K!$\/ tours and thus select the one that provides the best value for the $f$\/ function.

On the other hand, our tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use serial and multiprocessing heuristics to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}.

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $l_k$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n_k} t^k_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$\/ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section (Algorithm \ref{alg:itemsgen}).

@@

\vspace{2.0mm}
\begin{table}[H]
	\centering
	\caption{Testing scenarios}  \label{tab:scenarios}
	\begin{tabular}{c c c c }
		\toprule
		{\bf Scenario} & {$K$} & {$L$} & {\bf Aircraft} \\		
		\midrule
		1 & 2    & \{$l_0$, $l_1$, $l_2$\}                                 & smaller \\
		2 & 2    & \{$l_0$, $l_1$, $l_2$\}                                 & larger  \\
		3 & 3    & \{$l_0$, $l_1$, $l_2$, $l_3$\}                          & larger  \\
		4 & 4    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$\}                   & larger  \\
		5 & 5    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$, $l_5$\}            & larger  \\
		6 & 6    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$, $l_5$, $l_6$\}     & larger  \\
		\bottomrule
	\end{tabular}
\end{table}

@@

$UpdateConsolidated$, described in Algorithm \ref{alg:cons}, finds the best allocation for the consolidated items that remain on board. Initially, the set $Q$\/ is created, with the consolidated items that did not reach their destination (lines \ref{cons:remain1}-\ref{cons:remain2}). Then $MinCGDeviation$\/ (line \ref{cons:minCG}) is run through a MIP solver with the equations \ref{eq:torque}, \ref{eq:embarked} and \ref {eq:one}, to relocate the consolidated items on the pallets minimizing torque and ensuring that they all remain on board, one on each pallet. As there are few variables, the MIP solver returns an allocation $E^Q_k$\/ very quickly. Finally, the destination of each pallet with consolidated items is updated (lines \ref{cons:Ybegin}-\ref{cons:Yend}).

\begin{algorithm}[H]
	\caption{$UpdateConsolidated(k)$}  \label{alg:cons}
	\begin{algorithmic}[1]

		\State Let $Q_k  = \{ a^k_1, a^k_2, \ldots, a^k_{m_k} \}$ 
		\State $Q \gets \varnothing$ \label{cons:remain1}
		\For{$q\gets1$ to $m_k$} 
			\If {$a_q^k.to \in L_k$} 
				\State $Q \gets Q \cup \{a_q^k\}$
			\EndIf
		\EndFor	 
		\State $m_k \gets |Q|$ \label{cons:remain2}
		\State $E^Q_k \gets MinCGDeviation(k, Q)$ \label{cons:minCG}
		\For{$i \gets 1$ to $m$} \label{cons:Ybegin}
			\For{$q \gets 1$ to $m_k$}
				\If{$(p_i, a_q^k) \in E^Q_k$} 
					\State $p_i.to[k] \gets a_q^k.to$
				\EndIf
			\EndFor		
		\EndFor \label{cons:Yend}
		\State ${\bf return}\ Q, E^Q_k, M$ 
	\end{algorithmic}
\end{algorithm}

\begin{equation} \label{eq:torque}
	\mbox{minimize}\ | \sum_{i=1}^{m} \sum_{q=1}^{m_k} Y^k_{iq} \times p_i.d \times a_q^k.w |,\ \mbox{where}\ Q  = \{ a^k_1, a^k_2, \ldots, a^k_{m_k} \} 
\end{equation}

\begin{equation} \label{eq:embarked}
	s.t.: \sum_{i=1}^{m} Y^k_{iq} = 1;\ q \in \{1,2,\ldots,m_k\}
\end{equation}

\begin{equation} \label{eq:one}
	s.t.: \sum_{q=1}^{m_k} Y^k_{iq} \leq 1;\ i \in \{1,2,\ldots,m\}
\end{equation}

$SetPalletsDestination$, which set pallets destination not yet defined, is described in Algorithm \ref{alg:dest}. The vectors $vol[0..K]$\/ and $PalCons[0..K]$\/ store the total volume of items and the number of pallets with consolidated items destined for each node, respectively (lines \ref{dest:vector1}-\ref{dest:vector2}). The destinations of each pallet are defined proportionally to the volume of items, regarding the pallets with consolidated items (lines \ref{dest:propor1}-\ref{dest:propor2}). The node with the maximum volume demand defines the destination of any remaining pallets (lines \ref{dest:max1}-\ref{dest:max2}).

\begin{algorithm}[H]
	\caption{$SetPalletsDestination(k)$}  \label{alg:dest}
	\begin{algorithmic}[1]

		\For{$x \gets 0$ to $K$} \label{dest:vector1}
			\State $vol[x] \gets 0$ 
			\State $PalCons[x]  \gets 0$
		\EndFor
		\State $max \gets 0$ \label{dest:max}
		\State $total \gets 0$

		\For{$j \gets 1$ to $n_k$}
			\State $d \gets t_j^k.to$
			\If {$d \in L_k$} 
				\State $vol[d] \gets vol[d] + t_j^k.v$ 
				\State $total \gets total + t_j^k.v$ 
				\If {$vol[d] > vol[max]$}
					\State $max \gets d$ 
				\EndIf
			\EndIf
		\EndFor
		\For{$q \gets 1$ to $m_k$}
			\State $d \gets a_q^k.to$
			\State $PalCons[d] \gets PalCons[d]+1$
			\If {$d \in L_k$} 
				\State $vol[d] \gets vol[d] + a_q^k.v$
				\State $total \gets total + a_q^k.v$ 
				\If {$vol[d] > vol[max]$}
					\State $max \gets d$ 
				\EndIf
			\EndIf
		\EndFor \label{dest:vector2}
		\For{$x \gets 0$ to $K$} \label{dest:propor1}
			\If {$vol[x] \neq 0$}
				\State $quant \gets \max \{1, \lfloor{ m \times vol[x]/total}\rfloor \}-PalCons[x]$ 
				\State $np \gets 0$
				\For{$i \gets 1$ to $m$}

					\If{$np = quant$}
						\State ${\bf break}$
					\EndIf
					\If{$p_i.to[k] = -1$}
						\State $p_i.to[x] \gets x$
						\State $np \gets np + 1$
					\EndIf
				\EndFor

			\EndIf 
		\EndFor \label{dest:propor2}
		\For{$i \gets 1$ to $m$} \label{dest:max1}
			\If{$p_i.to[k] = -1$}
				\State $p_i.to[x] \gets max$
			\EndIf
		\EndFor \label{dest:max2}
		\State ${\bf return}\ M$

$ScoreAndDeviation$\/ is described in Algorithm \ref{alg:eval}, which evaluates the allocation graph generated by $SolveNode$, returning the corresponding score and CG deviation. It consists of a loop that goes through all the pallets (lines \ref{eval:loop1}-\ref{eval:loop2}), accumulating the scores (lines \ref{eval:score1}-\ref{eval:score2} ) and the torques (lines \ref{eval:eps1}-\ref{eval:eps2}) of the shipped items, allowing the final calculation of the CG deviation (lines \ref{eval:torque1}-\ref{eval:torque2}).

\begin{algorithm}[H]
	\caption{$ScoreAndDeviation(k, G)$}  \label{alg:eval}

	\begin{algorithmic}[1]
		\State Let $G(V_k, E^N_k \cup E^Q_k)$
		\State Let $X_{ij}^k = 1$ if $(p_i, t_j^k) \in E^N_k$, $1 \leq i \leq m$, $1 \leq j \leq n_k$
		\State Let $Y_{iq}^k = 1$ if $(p_i, a_q^k) \in E^Q_k$, $1 \leq i \leq m$, $1 \leq q \leq m_k$
		\State $s \gets 0$
		\State $\tau \gets 0$
		\State $weight \gets 0$
		\For{$i \gets 1$ to $m$}	\label{eval:loop1}
			\State $weight \gets weight + p_i.w$		
			\For{$j \gets 1$ to $n_k$}
				\If{$X_{ij}^k = 1$}
					\State $s \gets s + t_{ij}^k.s$ \label{eval:score1}
					\State $\tau \gets \tau + t_{ij}^k.w * p_i.d$ \label{eval:eps1}
				\EndIf
			\EndFor				
			\For{$q \gets 1$ to $m_k$}
				\If{$Y_{iq}^k = 1$}
					\State $s \gets s + a_{iq}^k.s$   \label{eval:score2}
					\State $\tau \gets \tau + a_{iq}^k.w * p_i.d$  \label{eval:eps2}
				\EndIf
			\EndFor	   \label{eval:loop2}
		\EndFor
		\State $maxWeight \gets \min(Payload, weight)$ \label{eval:torque1}
		\State $maxTorque \gets maxWeight \times limit^{CG}_{long}$
		\State $\epsilon \gets \tau/maxTorque$ \label{eval:torque2}
		\State ${\bf return}\ s, \epsilon$	
	\end{algorithmic}
\end{algorithm}

@@

The selection of edges for $E^N_k$\/ uses the {\it edge attractiveness}\/ $\theta^k_{ij}$\/ \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^k_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely proportional to the volume and the torque of each item. The 3000 value was empirically set to keep the attractiveness in the range [0, 1], to be compatible with the Ant Colony Optimization pheromone range.

>>>>

\usepackage{float}% http://ctan.org/pkg/float

@@

Transport aviation faces risks of cargo imbalance in the set of circumstances of aerial pickup and shipment of goods in a distribution system given the urgency required for loading for fast takeoff and mission accomplishment, notably in moments of crisis, public emergency, client contract short turnaround times, or any external forces for instant takeoff. Additionally, there is no computer support to aid load and trip programmers with the massive transportation demands in each location.

Other dangers are made possible as a result, including poor delivery, excessive fuel consumption, and longer turnaround times than necessary.

We developed an efficient solution by means of computer parallelism to the problem involving scheduling the loading and routing of an airplane in a tour of simultaneous pickup and delivery at intermediate hubs while taking into account a utility score, weight and balance rules, and fuel usage.

This hard problem, named {\it Air Cargo Load Planning with Routing, Pickup, and Delivery Problem} considers using standardized pallets in fixed positions, obeying the center of gravity constraints, delivering each item to its destination, and minimizing fuel consumption costs.

We also contributed by carrying out multiple experiments with the well-known Ant Colony Optimization on synthetic data based on real data from the {\it Brazilian Air Force}\/ transportation history and a new procedure to minimize the distance from the next node destined pallet to the ramp door.

We also created a process-based parallel computing heuristic that quickly finds good solutions for a wide range of problem sizes, an essential contribution as { \color{red} it was the unique method that managed to solve all testing scenarios.}

@@

Air cargo transport involves several sub-problems that are difficult to solve. Recently, \cite{MesquitaSanches2023} modeled and solved the {\it Air Cargo Load Planning with Routing, Pickup and Delivery} (ACLP+RPDP) composed by four sub-problems: {\it Build-up Scheduling Problem}, {\it Air Palletization Problem} (APP), {\it Weight and Balance Problem} (WBP), and a special case of the {\it Traveling Salesman Problem}, similar to the proposed by \cite{kaspi2019}, where the profit per unit time is defined as the profit gained after completing the tour divided by the total time (cost in our case) required to complete the tour.

However, there are still other important challenges in air cargo transport that go beyond the definition of the ACLP+RPDP, especially with regard to algorithms performance and the easiness of loading operations at each destination.

@@

		\cite{RoesenerHall2014}  & $\bigstar$ & $\bigstar$  & $.$  & $.$ & $.$ & $\bigstar$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{Vancroonemburg2014}  & $\bigstar$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{LurkinSchyns2015} & $.$ & $\bigstar$  & $\bigstar$ & $.$  & $.$ & $.$   & $.$  & $\bigstar$  &  $.$&  $.$ \\
		\cite{RoesenerBarnes2016}  & $.$ & $\bigstar$   & $.$ & $.$ & $.$ & $.$   & $\bigstar$  & $.$  &  $.$&  $.$ \\
		\cite{PaquaySchynsLimbourg2016,PaquayLimbourgSchynsOliveira2018}  & $\bigstar$ & $\bigstar$ & $.$ & $.$ & $.$ & $\bigstar$ & $\bigstar$  & $\bigstar$ & $.$&  $.$ \\
		\cite{YangLiuGao2018} & $.$ & $\bigstar$  & $.$  & $.$ & $\bigstar$  & $.$ & $\bigstar$ & $.$  & $.$&  $.$ \\
		\cite{wong2020} & $\bigstar$  & $\bigstar$  & $.$  & $.$   & $.$  & $.$ & $.$ & $\bigstar$  & $.$&  $.$  \\
		\cite{eugene2021} & $\bigstar$ & $\bigstar$ & $.$  & $.$   & $.$ & $.$ & $.$ & $\bigstar$  & $.$&  $.$  \\
		\cite{zhao2021} & $.$ & $\bigstar$ & $.$  & $.$  & $.$ & $.$ & $.$  & $\bigstar$ &  $.$&  $.$ \\

		\cite{MesquitaSanches2023}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$&  $.$  \\

		{\bf This work}   & $\bigstar$ & $\bigstar$  & $\bigstar$& $\bigstar$ & $.$ & $.$ & $\bigstar$ & $\bigstar$   &  $.$&  $\bigstar$  \\
		\bottomrule 
	\end{tabular}
	\normalsize 
\end{table}

@@

\cite{MesquitaSanches2023} proposed a method that attends these objectives: a pallet building and arrangement plan with a routing option that maximizes the benefit-cost ratio for the smooth execution of pickup and delivery transport missions. They develop a heuristic that can be executed on a simple handheld computer (like a laptop or a tablet) and that provides a solution quickly enough to keep this cargo handling time under an hour. This work seeks to improve their results in terms of quality and performance by the use of computer parallelism and by minimizing the distance between the next node destined pallets and the cargo door.

These improvements in their heuristics are meant to reduce the stress that transport planners are subjected to, because they have to deal with a lot of information in planning the aircraft route, assembling the pallets, and picking up and delivering at each node. To the best of our knowledge, that is the first time that an air cargo transport problem that simultaneously involves APP, WBP, PDP and TSP has been addressed, and the first time ACLP+RPDP is solved with computer parallelism.

@@

\cite{MongeauBes2003} presented a method based on linear integer programming to solve the problem of choosing and positioning containers on the {\it Airbus 340-300}. Safety and stability constraints were considered, with the objective of minimizing fuel consumption.

@@

As already mentioned, \cite{LurkinSchyns2015} was the first work that simultaneously modeled WBP and SPDP in air cargo transport. The authors demonstrated that this problem is NP-hard and performed some experiments with real data, noting that their model offers better results than those obtained manually.

@@

\cite{PaquaySchynsLimbourg2016} presented a mathematical modeling to optimize the loading of heterogeneous 3D boxes on pallets with a truncated parallelepipeds format. Its objective is to maximise the volume used in containers, considering load balancing constraints, the presence of fragile items and the possibility of rotating these boxes. \cite{PaquayLimbourgSchynsOliveira2018} developed some heuristics to solve this problem.

@@

\cite{wong2020}A developed a mathematical model and a tool based on mixed integer programming for optimizing cargo in aircraft with different pallet configurations. Balance constraints and the presence of dangerous items were considered. \cite{eugene2021} integrated this tool to a digital simulation model, with a visualization and validation system, based on sensors that alert about load deviations.

@@

\cite{MesquitaSanches2023} contributed with a complex model and elaborate heuristics for the ACLP-RPDP that simultaneously solve 4 intractable sub-problems: APP, WBP, SPDP and TSP. They also compare the performances of four well known heuristics: {\it Ant Colony Optimization}, the {\it Noising Method Optimization}, the {\it Greedy Randomized Adaptive Search Procedure}, and {\it Tabu Search}. They also create a new heuristic called {\it Shims} which is fast and may be run in a handheld computer. They did not use computer parallelism to improve performance, nor tryed to minimize the distances from the next nodes destined pallets to the cargo ramp door.

As can be seen, except for \cite{MesquitaSanches2023}, the remaining works do not address air cargo palletization and load balancing with route optimization in a multi-leg flight plan, and none of them employ parallelization in their algorithms. This is the objective of our work: to reshape \cite{MesquitaSanches2023} solution process and algorithms to accommodate parallel features that can improve solution quality and performance, and also include a new feature that is to minimize the distances between the next node's destined pallets and the cargo ramp door.

@@

	\item Pallets with destinations set to the next node should be put as near as possible to the cargo ramp door.

@@

Throughout this text, we call a {\it consolidated item} a set of items of the same destination stacked on a pallet and covered with a restraining net. It is considered unique, having the same attributes of its components, whose values are the sum of individual scores, weights and volumes. See Figure \ref{fig:larger2}.
Consolidated items must stay on board until they reach their destination to maintain accuracy in pickup and delivery procedures.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.4]{Images/consolidated.jpg}
	\caption{Consolidated items on 463L pallets inside a {\it Boeing C-17}}
	\small\textsuperscript{Source: From Wikimedia Commons, the free media repository}
	\label{fig:larger2}
\end{figure}

@@

We consider real scenarios with a smaller or a larger aircraft with payloads of 26,000 kg or 75,000 kg respectively. Both layouts are represented in Figures \ref{fig:smaller} and \ref{fig:larger}, where the pallets are identified by $p_i$.

@@

In both cases, the torque applied to the aircraft must keep its CG in the operational range, which corresponds to a percentage of the {\it Mean Aerodynamic Chord} \footnote{Chord is the distance between the leading and trailing edges of the wing, measured parallel to the normal airflow over the wing \cite[p.18]{HoughtonCarpenter2003}. The average length of the chord is known as the {\it Mean Aerodynamic Chord} (MAC).}: $0.556m$ in the smaller aircraft and $1.17m$ in the larger one. See Figure \ref{fig:lateral}.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.24]{Images/lateral.png}
	\caption{Aircraft longitudinal cut showing red lines as pallets}
	%\small\textsuperscript{Source: the author}
	\label{fig:lateral}
\end{figure}

@@

We accept the same assumptions as stated by \cite{MesquitaSanches2023}:
\begin{itemize}
	\item on each pallet, the items are distributed in such a way that their CG coincides with the centroid of the pallet;
	\item the CG of the payload must be at a maximum longitudinal distance of $limit^{CG}_{long}$ from the CG of the aircraft;
	\item in the larger aircraft, the CG of the payload must be at a maximum lateral distance of $limit^{CG}_{lat}$ from the CG of the aircraft;
	\item in the larger aircraft, pallets are distributed in two identical rows (with odd and even indices, respectively), and their centroids are at a distance $d^{CG}_{pallet}$ from the center-line of the aircraft.
\end{itemize}

@@

\bgroup
\def\arraystretch{1.2}
\begin{table}[H]
	\centering
	\small
	\begin{tabular}{MT}
		&Informally, ACLP+RPDP can be summarized as follows:\\
		\midrule
		max &  (items score sum) / (tour cost) of picked up and delivered items on a tour  \\
		\midrule
s.t.    & Along a tour, the set of unvisited nodes is updated. \\
		& In each node, an item may be included in at most one pallet.\\		
		& In each node, consolidated items are composed of items with the same destination. \\
		& Weight, volume and score of a consolidated item are the corresponding sum of their components.\\
		& Consolidated items remain on board until their destinations.\\	
		& Consolidated items can only be included in the same pallet if their destinations are the same.\\
		& Only items destined for the remaining nodes can be loaded.  \\
		& The lateral and longitudinal torques must be within the operational range of the aircraft.\\
		& Weight and volume limitations of pallets must be respected.\\
		& The total weight must be less than the aircraft payload or the total pallet capacity, whichever is the lowest.\\	
		& The pallets destined for the next node of the tour should be put as near as possible to the cargo ramp door (our inclusion).\\		
		\midrule
	\end{tabular}
	\normalsize
\end{table}
\egroup

@@

\cite{MesquitaSanches2023} strategy is based on the fact that, in real cases, $K$\/ is usually small. Specifically, they will consider $K \leq 6$\/ throughout their and our work, which is a higher value than usual in {\it Brazilian Air Force} missions. As a result, if we have fast node-by-node solutions that allow us to construct a complete tour, we will be able to test all possible $K!$\/ tours and thus select the one that provides the best value for the $f$\/ function.

The tactic will be, at each shipping node, to predefine the destinations of the pallets at that node. In this way, we will reserve a number of pallets proportional to the volume demanded by each destination at the shipping node. We could have used another criterion, but it was observed in the experiments that volume is more constrictive in airlift. Once the destinations of the pallets are defined, we will use serial and Multi-processing heuristics to find the best possible node-by-node solutions. This strategy is summarized in Algorithm \ref{alg:main}, retrieved from \cite{MesquitaSanches2023}.

@@

The other parameter $volume$\/ is a value greater than 1, which corresponds, at each node $l_k$, to the ratio between the sum of the volumes of the items ($\sum_{j=1}^{n_k} t^k_j.v$) and the load capacity of the pallets ($\sum_{i=1}^{m} p_i.v$). This parameter is passed to $ItemsGeneration$ (line \ref{main:items}), responsible for creating the items to be shipped, which will be presented in the next section.

@@

\vspace{2.0mm}
\begin{table}[H]
	\centering
	\caption{Testing scenarios retrieved from \cite{MesquitaSanches2023}.}  \label{tab:scenarios}
	\begin{tabular}{c c c c }
		\toprule
		{\bf Scenario} & {$K$} & {$L$} & {\bf Aircraft} \\		
		\midrule
		1 & 2    & \{$l_0$, $l_1$, $l_2$\}                                 & smaller \\
		2 & 2    & \{$l_0$, $l_1$, $l_2$\}                                 & larger  \\
		3 & 3    & \{$l_0$, $l_1$, $l_2$, $l_3$\}                          & larger  \\
		4 & 4    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$\}                   & larger  \\
		5 & 5    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$, $l_5$\}            & larger  \\
		6 & 6    & \{$l_0$, $l_1$, $l_2$, $l_3$, $l_4$, $l_5$, $l_6$\}     & larger  \\
		\bottomrule
	\end{tabular}
\end{table}

@@

$UpdateConsolidated$ finds the best allocation for the consolidated items that remain on board. Initially, the set $Q$ is created, with the consolidated items that did not reach their destination.

Then $MinCGDeviation$ is run through a MIP solver to relocate the consolidated items on the pallets minimizing torque and ensuring that they all remain on board, one on each pallet. As there are few variables, the MIP solver returns an allocation $E^Q_k$ very quickly.

$SetPalletsDestination$ sets pallets destination not yet defined. The total volume of items and the number of pallets with consolidated items destined for each node. The destinations of each pallet are defined proportionally to the volume of items, regarding the pallets with consolidated items.

$ScoreAndDeviation$ evaluates the allocation graph generated by $SolveNode$, returning the corresponding score and CG deviation. It consists of a loop that goes through all the pallets, accumulating the scores and the torques of the shipped items, allowing the final calculation of the CG deviation.

$UpdateConsolidated$, $MinCGDeviation$, $SetPalletsDestination$, and $ScoreAndDeviation$ are thoroughly described by \cite{MesquitaSanches2023}.

@@

The selection of edges for $E^N_k$\/ uses the {\it edge attractiveness}\/ $\theta^k_{ij}$, Equation \ref{eq:edge}, which can be understood as the tendency to allocate the item $t^k_j$\/ to the pallet $p_i$. It is directly proportional to the score, and inversely proportional to the volume and the torque of each item. The 3000 value was empirically set to keep the attractiveness in the range [0, 1], to be in the same order of magnitude of the Ant Colony Optimization pheromone range.
